---
title: "8_Predictions"
author: "Mia Akbar"
date: "`r Sys.Date()`"
output: html_document
---

# Library

```{r}
library(ggplot2)
library(gridExtra)
library(viridis)
```

Optional code to reduce clutter:

```{r}
source("http://bit.ly/theme_pub")
theme_set(theme_pub()) # Custom plotting theme
knitr::opts_chunk$set(warning = FALSE, message = FALSE) # Remove warnings about missing data
```

# Source

Model taken from JP09: MC Jones and A Pewsey A (2009) Sinh-arcsinh distributions. *Biometrika* 96: 761–780. 

As implemented as `sash` in the `GAMLSS` package:

$$f(y|\mu,\sigma,\nu,\tau)= \frac{\tau c}{\sigma \sqrt{2 \pi} (1+z^2)^{1/2}} \exp{(-\frac{1}{2} r^2)}$$

JP09 paper link: https://doi.org/10.1093/biomet/asp053

The probability density function is given in Eq 2 in JP09:

$$\frac{\delta C_{\epsilon,\delta}(x)}{\sqrt{2\pi(1+x^2)}}  e^{\frac{-S^2_{\epsilon,\delta}(x)}{2}}  $$

where 

$$C_{\epsilon,\delta}(x) = 
\cosh[\delta \sinh^{-1}(x)-\epsilon] = 
\sqrt{1+S^2_{\epsilon,\delta}(x)}$$

Solving for $S^2_{\epsilon,\delta}$:

$$S^2_{\epsilon,\delta}(x)=[\cosh(\delta \sinh^{-1}(x)-\epsilon)]^2-1$$

# Mean & Variance

To incorporate parameters for the mean ($\mu$) and variance ($\sigma$), we can simply substitute $x$ for $\frac{x-\mu}{\sigma}$ , yielding the full equation

$$\frac{\cosh[\delta \sinh^{-1}(\frac{x-\mu}{\sigma})-\epsilon]\delta}{\sqrt{(2\pi(1+(\frac{(x-\mu)^2}{\sigma^2})}}  \times exp{\frac{[\cosh(\delta \sinh^{-1}(\frac{x-\mu}{\sigma})-\epsilon)]^2-1}{2}}  $$

## Parameters

Example graphs are shown below, but in general:

  * $\mu$ = peak flowering
  * $\sigma$ = flowering variance
  * $\epsilon$ = skew paramater
  * $\delta$ = kurtosis parameter
  
To help understand these parameters, the graphs below show examples with different parameter values.

This is relatively easy to implement as a function in R:

```{r}
Nt<-function(t,m=0,s=1,e=0,d=1){
  if(d < 0 | s < 0){
    stop("delta (d) and sigma (s) must be > 0")
    } else {
      t<-(t-m)/s
    Cx<-cosh(d*asinh(t)-e)
    Sx<-(Cx^2-1)^(1/2)
    Nt<-(2*pi*(1+t^2))^(-1/2)*
      d*Cx*
      exp(-Sx^2/2)

    return(Nt/sum(Nt))
    }
}
```



# Plot

```{r}
t<-c(-1000:1000)/100
PDat<-data.frame(t=t,
                 Mid=Nt(t),
                 SkewL=Nt(t,e=-1,d=1),
                 SkewR=Nt(t,e=1,d=1),
                 Lepto=Nt(t,e=0,d=0.5),
                 Platy=Nt(t,e=0,d=2))
```

```{r}
Norm<-ggplot(aes(x=t,y=Norm),data=PDat) + 
  geom_line(colour="grey50", linewidth = 1, linetype="dotted") + 
  ylab("Nt")
```

## Predicted differences between latitudes  

### North 

Early start?
Early mean
Low variance
Positive skew
No adjustment to kurtosis parameter

```{r}
t = c(-1000:250)/100
NDat <- data.frame(t=t,
                 Pflwr = Nt(t, m = -7.5, s=0.8, e = 0.75, d = 0.5),
                 Lat = "North")
```

### Mid 

Later start?
Later mean
Higher variance
No adjustment in skew/kurtosis parameters

```{r}
t = c(-750:500)/100
MDat <- data.frame(t=t,
                 Pflwr = Nt(t, m = -0.75),
                 Lat = "Mid")
```


### South 

Later start?
Later mean
Higher variance
Negative skew
Leptokurtic

```{r}
t = c(-500:1000)/100
SDat <- data.frame(t=t,
                 Pflwr = Nt(t, m=5, s=2, e=-1, d = 1.5),
                 Lat = "South")
```

### Figure 2

```{r}
Dat <- rbind(NDat,MDat,SDat)
```


```{r}
PredFig <- ggplot(aes(x = t, y = Pflwr, colour = Lat), data = Dat) +
  geom_line(aes(group=Lat), linewidth = 1) +
  scale_color_manual(values = c("North" = "#150e38", "Mid" = "#ab337c", "South" = "#feb47b")) +
  ylab(expression(P[t])) +
  theme(axis.title.y = element_text(angle = 0),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank())
```

```{r export}
pdf("./Figures/Fig2.pdf",width=10,height=6)
PredFig
dev.off()
```
